### 2021-04-08 submission/manuscript.tex
Test updated manuscript after merging conflict

Sarah Tomkovich\^{}1, Ana Taylor, Jacob King, Joanna Colovas, Lucas
Bishop, Kathryn McBride, Sonya Royzenblat, Nicholas A. Lesniak, Ingrid
L. Bergin\^{}2, Patrick D. Schloss\textsuperscript{1\(\dagger\)}

@@

2. The Unit for Laboratory Animal Medicine, University off Michigan, Ann
Arbor, MI, USA

@@

\begin{itemize}
\item
  Medications have been shown to alter the intestinal microbiota, but
  how this may influence susceptibility to enteric infections is less
  clear.
\item
  The ubiquitous osmotic laxative, polyethylene glycol (PEG) 3350
  (Miralax/Nulytely/Golytely) that is also commonly used as bowel
  preparation for colonoscopies (also one of the possible administration
  routes used for fecal microbiota transplantation, a therapy for
  recurrent C. difficile infection) has been shown to influence the
  intestinal microbiota and mucosal barrier.
\item
  Review of previous mouse studies with PEG
\item
  2 main questions:

@@

@@

\#\#Discussion

\begin{itemize}
\item
  Summary of major findings
\item
  Discussion of prolonged persistence. C. difficile sequences detected
  in tissue samples. Association with mucin-degrading bacteria suggested
  by recent {[}a{[}ers{]}{]}
\item
  Discuss why we might not have observed more severe histology in PEG
  mice relative to clindamycin-treated mice

@@

\hypertarget{references}{%
\subsection{References}\label{references}}

>>>>

Sarah Tomkovich\^{}1, Ana Taylor, Jacob King, Joanna Colovas, Lucas
Bishop, Kathryn McBride, Sonya Royzenblat, Nicholas A. Lesniak, Ingrid
L. Bergin\^{}2, Patrick D. Schloss\textsuperscript{1\(\dagger\)}\$

@@

2. The Unit for Laboratory Animal Medicine, University of Michigan, Ann
Arbor, MI, USA

@@

Antibiotics are a major risk factor for \emph{Clostridioides difficile}
infections (CDIs) because of their impact on the microbiota. However,
antibiotics are not the only types of medications that disrupt the
microbiota. Although, other medications such as proton pump inhibitors
have been implicated as risk factors for CDIs through epidemiological
studies, whether the association is due to their impact on the
microbiome is still unclear (1). Many of the non-antibiotic medications
associated with CDIs are known to modulate intestinal motility, which in
turn also strongly impacts microbiota composition and function.\\
Interestingly our group has previously demonstrated that there are
similar microbiota features in stool samples from CDI patients and
patients with diarrhea that tested negative for \emph{C. difficile} with
both groups having lower alpha diversity than non-diarrheal control
samples (2). These results led to our hypothesis that bacterial
communities from patients experiencing diarrheal controls are
susceptible, but have not been exposed to \emph{C. difficile} spores. To
examine if diarrhea alone is sufficient to disrupt microbiota
colonization resistance, we turned to osmotic laxatives, which increase
intestinal motility and can lead to diarrhea. The ubiquitous osmotic
laxative, polyethylene glycol (PEG) 3350 is found in Miralax, Nulytely,
and Golytely and is also commonly used as bowel preparation for
colonoscopies. Interestingly, previous studies has shown that treating
mice with PEG alone rendered the mice susceptible to \emph{C. difficile
}infection, altered microbiota composition, reduced acetate and butyrate
and altered the mucus barrier (3--6). The mucus barrier is thought to
mediate protection from \emph{C. difficile} infections by protecting
intestinal epithelial cells from the toxins produced by \emph{C.
difficile} (Ref). PEG is also relevant in the context of treating
recurrent CDIs via fecal microbiota transplant (FMT). For FMTs that are
delivered via colonoscopy, patients typically undergo bowel preparation
by taking an osmotic laxative prior to the procedure. Many of the FMT
studies to date rationalize the use of laxatives (Ref) based on a case
study with 2 pediatric patients where the authors suggested the laxative
may help flush \emph{C. difficile} spores and toxins from the intestine
(7).

\begin{itemize}
\tightlist
\item
  2 main questions:

@@

\textbf{Laxative treatment alone leads to prolonged \emph{C. difficile}
colonization in mice.} We compared PEG-treated mice to our standard 10
mg/kg clindamycin treatment, which temporarily renders the mice
susceptible to \emph{C. difficile}, with mice typically clearing C.
difficile within 10 days post-infection (1, 8). All PEG-treated mice
were administered a 15\% PEG solution in the drinking water for 5-days,
one group was also treated with clindamycin, and one group was allowed
to recover for 10 days prior to challenge (Fig. 1A). After PEG and/or
antibiotic treatment all mice were challenged with 10\textsuperscript{3}
\emph{C. difficile} 630 spores.

@@

\hypertarget{discussion}{%
\subsection{Discussion}\label{discussion}}

\begin{itemize}
\item
  Summary of major findings
\item
  Discussion of prolonged persistence. C. difficile sequences detected
  in tissue samples. Association with mucin-degrading bacteria suggested
  by recent papers.
\item
  Discuss why we might not have observed more severe histology in PEG
  mice relative to clindamycin-treated mice

@@

\hypertarget{references}{%
\subsection*{References}\label{references}}
\addcontentsline{toc}{subsection}{References}

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-Tomkovich2019}{}%
1. \textbf{Tomkovich S}, \textbf{Lesniak NA}, \textbf{Li Y},
\textbf{Bishop L}, \textbf{Fitzgerald MJ}, \textbf{Schloss PD}. 2019.
The proton pump inhibitor omeprazole does not promote
\emph{Clostridioides difficile} colonization in a murine model. mSphere
\textbf{4}.
doi:\href{https://doi.org/10.1128/msphere.00693-19}{10.1128/msphere.00693-19}.

\leavevmode\hypertarget{ref-Schubert2015}{}%
2. \textbf{Schubert AM}, \textbf{Sinani H}, \textbf{Schloss PD}. 2015.
Antibiotic-induced alterations of the murine gut microbiota and
subsequent effects on colonization resistance against \emph{Clostridium
difficile}. mBio \textbf{6}.
doi:\href{https://doi.org/10.1128/mbio.00974-15}{10.1128/mbio.00974-15}.

\leavevmode\hypertarget{ref-Kashyap2013}{}%
3. \textbf{Kashyap PC}, \textbf{Marcobal A}, \textbf{Ursell LK},
\textbf{Larauche M}, \textbf{Duboc H}, \textbf{Earle KA},
\textbf{Sonnenburg ED}, \textbf{Ferreyra JA}, \textbf{Higginbottom SK},
\textbf{Million M}, \textbf{Tache Y}, \textbf{Pasricha PJ},
\textbf{Knight R}, \textbf{Farrugia G}, \textbf{Sonnenburg JL}. 2013.
Complex interactions among diet, gastrointestinal transit, and gut
microbiota in humanized mice. Gastroenterology \textbf{144}:967--977.
doi:\href{https://doi.org/10.1053/j.gastro.2013.01.047}{10.1053/j.gastro.2013.01.047}.

\leavevmode\hypertarget{ref-Ferreyra2014}{}%
4. \textbf{Ferreyra JA}, \textbf{Wu KJ}, \textbf{Hryckowian AJ},
\textbf{Bouley DM}, \textbf{Weimer BC}, \textbf{Sonnenburg JL}. 2014.
Gut microbiota-produced succinate promotes c.~difficile infection after
antibiotic treatment or motility disturbance. Cell Host \& Microbe
\textbf{16}:770--777.
doi:\href{https://doi.org/10.1016/j.chom.2014.11.003}{10.1016/j.chom.2014.11.003}.

\leavevmode\hypertarget{ref-Tropini2018}{}%
5. \textbf{Tropini C}, \textbf{Moss EL}, \textbf{Merrill BD}, \textbf{Ng
KM}, \textbf{Higginbottom SK}, \textbf{Casavant EP}, \textbf{Gonzalez
CG}, \textbf{Fremin B}, \textbf{Bouley DM}, \textbf{Elias JE},
\textbf{Bhatt AS}, \textbf{Huang KC}, \textbf{Sonnenburg JL}. 2018.
Transient osmotic perturbation causes long-term alteration to the gut
microbiota. Cell \textbf{173}:1742--1754.e17.
doi:\href{https://doi.org/10.1016/j.cell.2018.05.008}{10.1016/j.cell.2018.05.008}.

\leavevmode\hypertarget{ref-VanInsberghe2020}{}%
6. \textbf{VanInsberghe D}, \textbf{Elsherbini JA}, \textbf{Varian B},
\textbf{Poutahidis T}, \textbf{Erdman S}, \textbf{Polz MF}. 2020.
Diarrhoeal events can trigger long-term clostridium difficile
colonization with recurrent blooms. Nature Microbiology
\textbf{5}:642--650.
doi:\href{https://doi.org/10.1038/s41564-020-0668-2}{10.1038/s41564-020-0668-2}.

\leavevmode\hypertarget{ref-Liacouras1996}{}%
7. \textbf{Liacouras CA}, \textbf{Piccoli DA}. 1996. Whole-bowel
irrigation as an adjunct to the treatment of chronic, relapsing
clostridium difficile colitis. Journal of Clinical Gastroenterology
\textbf{22}:186--189.
doi:\href{https://doi.org/10.1097/00004836-199604000-00007}{10.1097/00004836-199604000-00007}.

\leavevmode\hypertarget{ref-Tomkovich2020}{}%
8. \textbf{Tomkovich S}, \textbf{Stough JMA}, \textbf{Bishop L},
\textbf{Schloss PD}. 2020. The initial gut microbiota and response to
antibiotic perturbation influence clostridioides difficile clearance in
mice. mSphere \textbf{5}.
doi:\href{https://doi.org/10.1128/msphere.00869-20}{10.1128/msphere.00869-20}.
